---
title: "ESM 206 Lab 6"
author: "Kirsten White"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)

```

## Example of a rank-based test 
Using our own samples using a pseudorandom generator (with set.seed(insert_number))
Set.seed tells r to create vectors of random samples -- 
We will be comparing rank across groups of samples instead of values.

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE) # To create sample of integer values, inside parenthesis tell it range to select integers from (0-20),  and how many values in it (size = 15), then tell it to replace (can see same integer values more than once in your set; value gets replaced into the pool of integers to choose from after used once)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

### Let's compare these samples using histograms
```{r}
hist(gp_1)

hist(gp_2)
```
### Try a t-test
Use this when the mean is the appropriate measure of standard deviation. 

```{r}
t.test(gp_1, gp_2)
```

This p-value means: if these samples were drawn from populations with the same mean (the null hypothesis), the probability of taking 2 random samples with means *at least as different* as the sample means we found by random chance (accounting for spread and sample size) is 19.82%. 

Given that this is pretty likely to occur, we will *retain the null hypothesis*. (Fail to reject the null hypothesis -> no significant difference between group 1 and group 2)

Warning: people get weirdl upset if you say "accept" the null. 

Now, let's compare this outcome to a rank-based test. 

### Mann Whitney U unpaired rank-based test
Use this when medians are the appropriate measure to compare the data (standard deviation).
Learn more on this by typing ?wilcox.test into the console. See what outputs mean by scrolling all the way down to the values section that describes each. 

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu
```
What does this outcome actually mean (p-value of 0.28)? 

If these samples were drawn from populations with the same ranks (medians), the probability of finding 2  samples with ranks *at least as different* as those in our samples by random chance is 0.28 (28%). This probability is pretty likely to occur, so we do not have enough evidence to reject the null hypothesis; therfore, we will retain the null hypothesis (plausible these values were pulled from populations with the same ranks). 

There is no significant difference in ranks between group 1 and group 2. Then, (provide statistical summary here).  



Now, let's get into linear regression!

## Linear Regression 
Simple linear regression (single dependent variable, a single independent variable)

Looking at penguins data, use penguin flipper length v penguin body mass

```{r, include = FALSE}
# Make an exploratory plot of penguin body mass (y axis) versus flipper length (x axis) 

ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g))+
  geom_point() +
  geom_smooth(method = "lm")
```

Now, find a linear regression model using ordinary least squares describing the relationship between flipper length and body mass for these penguins. 

3 pieces: 
- What type of model? 
- What is the relationship to model (DV - IV(s))? 
- Where is the data that is used to create this model?

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
```
What does this mean? missed this part of lecture bc my code wasnt working

interpret coeff as expected avg change in...

What is the actual interpretation of intercept (value of -5780.83)? -> if penguin with flipper length = 0 mm then a body mass of that penguin would be -5780.83

If you just want the coeff associated with flipper length -> type penguin_lm$coefficients[2] in the console to pull it out 

### Broom packages returns model outputs as tidy dataframes 
Can use broom to access individual pieces - shows st error, statistic, p value 
Lots of useful information included in here that you might want to use in your table 
using broom::glance(dataframe) shows model wide 
```{r}
penguin_lm_tidy <- broom::tidy(penguin_lm) # Can also just use tidy()
penguin_lm_tidy
```

How can I actually include my model equation in a report?

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE) # use_coefs replaces with the actual model coefs that were calculated
```
```{r}
plot(penguin_lm)
```
Compares each of the observed values to what the model predicts it should be 
1) Residual v fitted - how far - in this case, looks relatively even spread around that predictive line 
3) scale - location - does it look like the residuals have equal variance across the predictive model?

diagnostic looking at residuals distribution 
2) normal QQ - explores normality, but here normality is referencing the value of the residuals (assumption is normality of the residuals)- in this case, residuals look very very normally distributed 

measure of leverage of any point; measures outliers; what points seem to be most impacting the (only remove outliers if we think the data point is not actually representative of our population, NOT just because its different)
4) residuals v leverage 