---
title: "ESM 206 Lab 6"
author: "Kirsten White"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)

```

## Example of a rank-based test 
Using our own samples using a pseudorandom generator (with set.seed(insert_number))
Set.seed tells r to create vectors of random samples -- 
We will be comparing rank across groups of samples instead of values.

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE) # To create sample of integer values, inside parenthesis tell it range to select integers from (0-20),  and how many values in it (size = 15), then tell it to replace (can see same integer values more than once in your set; value gets replaced into the pool of integers to choose from after used once)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

### Let's compare these samples using histograms
```{r}
hist(gp_1)

hist(gp_2)
```
### Try a t-test
Use this when the mean is the appropriate measure of standard deviation. 

```{r}
t.test(gp_1, gp_2)
```

This p-value means: if these samples were drawn from populations with the same mean (the null hypothesis), the probability of taking 2 random samples with means *at least as different* as the sample means we found by random chance (accounting for spread and sample size) is 19.82%. 

Given that this is pretty likely to occur, we will *retain the null hypothesis*. (Fail to reject the null hypothesis -> no significant difference between group 1 and group 2)

Warning: people get weirdl upset if you say "accept" the null. 

Now, let's compare this outcome to a rank-based test. 

### Mann Whitney U unpaired rank-based test
Use this when medians are the appropriate measure to compare the data (standard deviation). 

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu
```
What does this outcome actually mean (p-value of 0.28)? 

If these samples were drawn from populations with the same ranks (medians), the probability of finding 2  samples with ranks *at least as different* as those in our samples by random chance is 0.28 (28%). This probability is pretty likely to occur, so we do not have enough evidence to reject the null hypothesis; therfore, we will retain the null hypothesis (plausible these values were pulled from populations with the same ranks). 

There is no significant difference in ranks between group 1 and group 2. Then, (provide statistical summary here).  
